 

### 2020-10-06 vertical-thermal-review.Rmd
incorporated some changes from group discussion

v.	What are the implications of these patterns for the ecology and climate change responses of canopy versus understory trees?

vi. What are the implications for modeling/ scaling?

@@

*(paragraph on leaf area distribution, possibly including figure? [see GitHub issue #15](https://github.com/EcoClimLab/vertical-thermal-review/issues/15))*
The vertical structure and hence, microenvironments of forest canopies vary seasonally (Parker & Tibbs 2004; Smith et al. 2019 ; Parker et al. 2019) and following disturbance (e.g., Parker et al. 2002 ; Almeida et al. 2016 ; Stark et al. 2020 in press).
In this review, we focus primarily on growing season conditions, including leaf phenology at the edges of the growing season but excluding periods (in deciduous forests) when no leaves are present.

Light conditions, specifically photosynthetically active radiation (PAR) strongly varies  with canopy height due to cumulative layers of vegetation that limits light to lower canopy layers [@fauset_differences_2018; @niinemetsPackingPhotosyntheticMachinery2009; Parker 1995; @poorterMetaanalysisPlantResponses2019b]
(Fig. **2x**)such that the forest floor of understory regions receive 1-2%  of the light incident to the top of the canopy [@robertsStomatalBoundaryLayerConductances1990a; Chazdon & Fetcher 1984; Parker et al. 2019]
In heterogeneous canopies [with high gap fractions], distance from the outer canopy is a better proxy for light environment than height (Parker 1995) and lower canopy light environments are highly variable, ranging from high light gap to deeply shaded understory regions [@chazdonPhotosyntheticLightEnvironments1984; @tymenQuantifyingMicroenvironmentalVariation2017; @smithSeasonalDroughtrelatedChanges2019]. 
This difference in light regime is an important variable in determining leaf traits along the gradient and consequently their responses to temperature [@fausetDifferencesLeafThermoregulation2018; @niinemetsResponsesForestTrees2010, @sackHowStrongIntracanopy2006; @doughtyAreTropicalForests2008].

- (*@bonan_ecological_2016 reviews this and points to appropriate references*)
-	Sunflecks: Leaky et al. 2003--from Marielle, "This is another, higher resolution axis of the variability of microenvironments that is probably quite important to address. (Also see the papers by Robin Chazdon on temporal variability of light environments). Maybe we will want to add these sentences (above) and something about sunflecks to a separate section that specifically addresses sources of variation in forest structure & microenvironments (e.g., vertical, horizontal, seasonal, diurnal, and with disturbance)?"

@@

## Leaf temperature

@@

@@

## Ecology

@@

@@

## Implications for Modeling / Scaling / Future Climate Projections

@@

#### Future Questions

*(KAT thinks this section should probably be dropped/ content integrated elsewhere, most within the implications for modeling/ scaling section)*

- Are patterns (in traits, metabolism, ecology) driven by tree height or exposure?

>>>>

v.	What are the implications of these patterns for the ecology and climate change responses of canopy versus understory trees? What are the implications for modeling/ scaling? (*nerge these better*)

@@

- (*@bonan_ecological_2016 reviews this and points to appropriate references*)

*(paragraph on leaf area distribution, including figure ([separate from or combined with current Fig. 2- see GitHub issue #15](https://github.com/EcoClimLab/vertical-thermal-review/issues/15))*
The vertical structure and hence, microenvironments of forest canopies vary seasonally (Parker & Tibbs 2004; Smith et al. 2019 ; Parker et al. 2019) and following disturbance (e.g., Parker et al. 2002 ; Almeida et al. 2016 ; Stark et al. 2020 in press).
In this review, we focus primarily on growing season conditions, including leaf phenology at the edges of the growing season but excluding periods (in deciduous forests) when no leaves are present.

Light conditions, specifically photosynthetically active radiation (PAR) strongly varies  with canopy height due to cumulative layers of vegetation that limits light to lower canopy layers [@fauset_differences_2018; @niinemetsPackingPhotosyntheticMachinery2009; Parker 1995; @poorterMetaanalysisPlantResponses2019b]
(Fig. **2x**)such that the forest floor of understory regions receive 1-2%  of the light incident to the top of the canopy [@robertsStomatalBoundaryLayerConductances1990a; Chazdon & Fetcher 1984; Parker et al. 2019]
In heterogeneous canopies [with high gap fractions], distance from the outer canopy is a better proxy for light environment than height (Parker 1995) and lower canopy light environments are highly variable, ranging from high light gap to deeply shaded understory regions [@chazdonPhotosyntheticLightEnvironments1984; @tymenQuantifyingMicroenvironmentalVariation2017; @smithSeasonalDroughtrelatedChanges2019]. 
This difference in light regime is an important variable in determining leaf traits along the gradient and consequently their responses to temperature [@fausetDifferencesLeafThermoregulation2018; @niinemetsResponsesForestTrees2010, @sackHowStrongIntracanopy2006; @doughtyAreTropicalForests2008].
-	Sunflecks: Leaky et al. 2003--from Marielle, "This is another, higher resolution axis of the variability of microenvironments that is probably quite important to address. (Also see the papers by Robin Chazdon on temporal variability of light environments). Maybe we will want to add these sentences (above) and something about sunflecks to a separate section that specifically addresses sources of variation in forest structure & microenvironments (e.g., vertical, horizontal, seasonal, diurnal, and with disturbance)?"

@@

- *Are patterns (in traits, metabolism, ecology) driven by tree height or exposure?*

## Leaf temperature and hydraulics

@@

*(Because hydraulics are key to leaf T, we should include some of that here--e.g., water on leaves and leaf water content)*

@@

## Scaling & implications

### Distribution of species and traits across the vertical profile

@@

### Scaling to tree level: demography

@@

### Representation in models

@@

*(do we need a section on ecosystem function-- e.g., transpiration, gpp?)*

### Implications for climate change responses
